### 2020-03-10 random/career_preparation/cover_letters/general/AQR.tex
updates

\documentclass[11pt,a4paper]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
\renewcommand*\familydefault{\sfdefault}

\usepackage{graphicx} % Required for including pictures
\begin{document}

Dear AQR,

I am interested in the Research Analyst position for Summer 2020.
I have experience managing cloud instances, SQL databases, and on-line machine learning.
My project to track sports data is reflective of this. 
Initially in python and dreadfully slow, I could only gather a few sports every couple of minutes.
After porting my code to Golang and concurrently requesting REST endpoints, I am able to track 2000+ sports in under 30 seconds.
Additionally, now that the script writes to a Cloud SQL database, instead of CSVs, it is easier to track neural network predictions from PyTorch.
I realize there is still much to improve with my project (submit a PR!), but I see my ability to identify bottlenecks and iterate as a great strength. 
I believe this project demonstrates why I am fit for this position.

I hope to hear from you.

Thank you,

Anand Jain\\
anandj@uchicago.edu\\
Github, Linkedin: anandijain

\end{document}