### 2020-03-10 random/career_preparation/cover_letters/general/betfair.tex

\documentclass[10pt,a4paper]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
\renewcommand*\familydefault{\sfdefault}

\usepackage{graphicx} % Required for including pictures
\begin{document}

Dear Betfair,

I am interested in a summer internship, although I do not have a visa to work in Dublin.

I have built a Go scraper for Bovada that runs concurrently and gets the scores and moneylines for 1600+ games in 25 seconds.

I check the changes from the previous heartbeat and write these to disk, for use in Pytorch training.

I have built Long Short Term Memory networks, 1D-convolutional networks, and reinforcement learning agents for this data and feel confident in implementing classical and neural network machine learning applications.

One thing that excites me about working at Betfair is the ability to analyze and understand exchanges and behavioral economics in a domain outside of traditional finance.

A specific question that I have is on whether or not liquidity is a financial analog that holds in sports exchanges.

I hope you consider me for a position this summer.

Thank you,

Anand Jain\\
anandj@uchicago.edu\\
Github, Linkedin: anandijain

\end{document}