### 2020-03-13 random/career_preparation/cover_letters/ND_paper.tex
update

\documentclass[11pt,a4paper]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
\renewcommand*\familydefault{\sfdefault}

\usepackage{graphicx} % Required for including pictures
\begin{document}

Dear Hiring Manager,

ND presents an interesting set of opportunities to apply innovative technologies to quantifiable and sustainable recycling.

These challenges are particularly interesting to me because I feel committed to a good future for our children and I am interested in computer vision and machine learning.

I have qualifications in applied machine learning which will help decrease the contamination rate in sorted goods and provide a more reusable and recyclable product.

One area that I am excited to learn more about, which this opportunity presents, is a hands on understanding of industrial optimization and robotics.

I have experience in Excel, but I am confident in my capabilities in Python’s Matplotlib, which is a robust data visualization package.

Thank you,

Anand Jain\\
anandj@uchicago.edu\\
Github, Linkedin: anandijain

\end{document}