### 2020-03-13 random/career_preparation/cover_letters/dolby.tex

\documentclass[11pt,a4paper,english]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
\renewcommand*\familydefault{\sfdefault}

\usepackage{graphicx} % Required for including pictures
\usepackage{babel}
\usepackage[useregional]{datetime2}
\begin{document}

Dear Dolby,

I am interested in the Machine Learning / Signal Processing Internship for Summer 2020.

As a lifelong musician and producer, this opportunity is extremely appealing to me.

I have extensive experience with git, python, and go (similar to C++), as well as using cloud APIs for databases, virtual machines, and flask applications.

I built a tool to generate white noise and apply filters to noise in C at /DSP/ and used a variational autoencoder to generate sound based off my personal songs at /sippyart/.

I see this opportunity as an exciting way to learn about low latency audio, networking, testing, and other interesting parts of large infrastructure that don't necessarily get taught in a class.

One area that I would like to learn more from a team about is how noise reduction is implemented and how hardware designs are constructed to optimally flatten the frequency response.

Thank you,

Anand Jain\\
anandj@uchicago.edu\\
Github, Linkedin: anandijain\\
\today\\

\end{document}