### 2020-03-13 random/career_preparation/cover_letters/nielsen.tex

\documentclass[11pt,a4paper,english]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
\renewcommand*\familydefault{\sfdefault}

\usepackage{graphicx} % Required for including pictures
\usepackage{babel}
\usepackage[useregional]{datetime2}
\begin{document}

Dear Nielsen,

I am interested in being a Data Science Internship for Summer 2020.

I have extensive experience with Python, Git, Google Cloud, and SQL.

I have built a Go-lang based data-scraper for sports odds that initially took 20+ minutes in python, now takes 30 seconds.

I am interested in how larger scale data operations work and how to find interesting correlations and abstractions in constrained and big-data scenarios.

I hope to hear from you.

Thank you,

Anand Jain\\
anandj@uchicago.edu\\
Github, Linkedin: anandijain\\
\today\\

\end{document}