### 2020-03-13 random/career_preparation/cover_letters/template.tex

\documentclass[11pt,a4paper,english]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
\renewcommand*\familydefault{\sfdefault}

\usepackage{graphicx} % Required for including pictures
\usepackage{babel}
\usepackage[useregional]{datetime2}
\usepackage{stix} % Use the Stix font by default

\usepackage{microtype} % Improve justification

\signature{Anand Jain} % Your name for the signature at the bottom

\address{610 Jackson St. \\ Santa Clara, CA 95050 \\ } % Your address and phone number

%----------------------------------------------------------------------------------------

\begin{document}

%----------------------------------------------------------------------------------------
%	ADDRESSEE SECTION
%----------------------------------------------------------------------------------------

\begin{letter}{Nathan Drake \\ LaTeXTemplates.com \\ 321 Pleasant Lane \\ City, State 12345} % Name/title of the addressee

%-------------------------

Dear ,

I am interested in your  position for Summer 2020.

Thank you for your time and consideration.

I look forward to your reply.

Thank you,

Anand Jain\\
anandj@uchicago.edu\\
Github, Linkedin: anandijain\\
\today\\
\end{letter}
\end{document}