### 2020-03-13 random/career_preparation/cover_letters/template2.tex

\documentclass[11pt,a4paper]{letter} % 10pt font size default, 11pt and 12pt are also possible
\renewcommand*\familydefault{\sfdefault}

\address{610 Jackson St. \\ Santa Clara, CA 95050 \\ (408)597-4214 } % Your address and phone number
% \address{6103 S University Ave. \\ Chicago, IL 60637 \\ (408)597-4214 } % Your address and phone number

\begin{document}

\begin{letter}{Company \\ Company Address \\ City, State 12345}

\opening{\textbf{Dear Sir or Madam,}}

Thank you for your time and consideration.

I look forward to your reply.

\closing{Sincerely, \\ Anand Jain \\ University of Chicago 2021 \\ anandj@uchicago.edu}

\end{letter}

\end{document}