### 2020-03-13 random/career_preparation/cover_letters/verily.tex

\documentclass[11pt,a4paper,english]{letter} % Specify the font size (10pt, 11pt and 12pt) and paper size (letterpaper, a4paper, etc)
\renewcommand*\familydefault{\sfdefault}

\usepackage{graphicx} % Required for including pictures
\usepackage{babel}
\usepackage[useregional]{datetime2}
\begin{document}

Anand Jain\\
anandj@uchicago.edu\\
Github, Linkedin: anandijain\\

Dear Verily,

I am interested in the Machine Learning/Software Engineering Internship for Summer 2020.

AlphaFold was one advance that sparked my interest in computational biology and made me think about how I could use my learning to have real impact.

In computer science, we learn about the layers of abstraction that are required to build up systems for high level programming.

Biology is incredibly interesting, because we have to decrypt eons of evolution's abstractions to understand the systems and use it to improve others' lives.

I have experience in writing machine learning applications that run on Google Cloud virtual machines using PyTorch, and locally using GPU.

I have experience with Go-lang and SQL for larger scale systems through my sports projects.

I am interested in contributing towards Project Baseline and learning more about biological modelling.

Thank you,

Anand Jain\\
\today\\

\end{document}