### 2020-03-16 random/career_preparation/cover_letters/impossible.tex

\documentclass[11pt,a4paper]{letter} % 10pt font size default, 11pt and 12pt are also possible
\renewcommand*\familydefault{\sfdefault}

\address{610 Jackson St. \\ Santa Clara, CA 95050 \\ (408)597-4214 } % Your address and phone number
% \address{6103 S University Ave. \\ Chicago, IL 60637 \\ (408)597-4214 } % Your address and phone number

\begin{document}

\begin{letter}{Impossible Foods \\ Redwood City, CA}

\opening{\textbf{Dear Impossible Foods,}}

I am interested in being a Data Science Intern for Summer 2020.

I believe I have the qualifications in programming, machine learning, and cloud service management to provide valuable insights on social media sentiment.
Much of my time in college has been devoted towards thinking about neural network architectures and paradigms, differentiable programming and reinforcement learning being two primary interests. 
I am looking to work on larger and more complex models, such as BERT and GPT-2.

I have experience working with the hugging-face NLP transformers library, and will be taking computational linguistics next term.
My ideal conversation with an employee would be about how simulation and machine learning can be used to continually improve the Impossible Burger.

Thank you for your time and consideration.

I look forward to your reply.

\closing{Sincerely, \\ Anand Jain \\ University of Chicago 2021 \\ anandj@uchicago.edu}

\end{letter}

\end{document}