### 2020-03-16 random/career_preparation/cover_letters/intel.tex

\documentclass[11pt,a4paper]{letter} % 10pt font size default, 11pt and 12pt are also possible
\renewcommand*\familydefault{\sfdefault}

\address{610 Jackson St. \\ Santa Clara, CA 95050 \\ (408)597-4214 } % Your address and phone number
% \address{6103 S University Ave. \\ Chicago, IL 60637 \\ (408)597-4214 } % Your address and phone number

\begin{document}

\begin{letter}{Intel Corporation \\ 2200 Mission College Blvd \\ Santa Clara, CA  95052 }

\opening{\textbf{Dear Intel,}}

In senior year of high school, I was taking AP Computer Science Principles and we had to do two presentations. I did my first presentation on convolutional neural networks and my second on quantum computing. 
I didn’t realize until I learned more about complexity theory and big-O notation in college how amazing quantum systems are. 
This profound and numinous field of study is racing to Landauer's limit and has given me an exponentially greater appreciation for the beauty of nature.

I'm hoping to contribute some meaningful addition to helping us use NISQ devices to improve lives around the world.
As founder of the UChicago Quantum Society, I am hoping to bring back insights in how to communicate scientific works to general audiences and help grow the community of the quantum-curious.

\closing{Sincerely, \\ Anand Jain \\ University of Chicago 2021 \\ anandj@uchicago.edu}

\end{letter}

\end{document}