### 2020-03-16 random/career_preparation/cover_letters/pg_and_e.tex

\documentclass[11pt,a4paper]{letter} % 10pt font size default, 11pt and 12pt are also possible
\renewcommand*\familydefault{\sfdefault}

\address{610 Jackson St. \\ Santa Clara, CA 95050 \\ (408)597-4214 } % Your address and phone number
% \address{6103 S University Ave. \\ Chicago, IL 60637 \\ (408)597-4214 } % Your address and phone number

\begin{document}

\begin{letter}{Pacific Gas and Electric \\ 77 Beale Street \\ San Francisco, CA 94105}

\opening{\textbf{Dear Pacific Gas and Electric,}}

I am interested in being an Meteorology Data Science Intern for Summer 2020.

I believe I have the qualifications in programming, machine learning, and deploying cloud services to make a helpful contribution towards creating innovative predictive models.
Much of my time in college has been devoted towards thinking about neural network architectures and paradigms, differentiable programming and reinforcement learning being two primary interests. 
I am looking to implement larger scale models that use interesting data and require innovative models, which meteorology requires.

My ideal conversation with an employee would be about what architectures and models are used currently in meteorology and what the constraints are of these models.

Thank you for your time and consideration.

I look forward to your reply.

\closing{Sincerely, \\ Anand Jain \\ University of Chicago 2021 \\ anandj@uchicago.edu}

\end{letter}

\end{document}