### 2020-03-16 random/career_preparation/cover_letters/pocket_gems.tex

\documentclass[11pt,a4paper]{letter} % 10pt font size default, 11pt and 12pt are also possible
\renewcommand*\familydefault{\sfdefault}

\address{610 Jackson St. \\ Santa Clara, CA 95050 \\ (408)597-4214 } % Your address and phone number
% \address{6103 S University Ave. \\ Chicago, IL 60637 \\ (408)597-4214 } % Your address and phone number

\begin{document}

\begin{letter}{Pocket Gems \\ 220 Montgomery Street \\ San Francisco, CA 94104}

\opening{\textbf{Dear Pocket Gems,}}

I am interested in being a Data Science Intern for Summer 2020.

I believe I have the qualifications in programming, machine learning, and cloud service management to develop and deploy valuable and performant models.
Much of my time in college has been devoted towards thinking about neural network architectures and paradigms, differentiable programming and reinforcement learning being two primary interests. 
I am looking to work on larger and more complex models.

I have experience working with the hugging-face NLP transformers library, and will be taking computational linguistics next term.
I have built reinforcement learning environements using OpenAI's Gym package and am excited to implement larger models in production.
My ideal conversation with an employee would be about how AI is changing the gaming industry and what creative uses of AI in gaming there are.

Thank you for your time and consideration.

I look forward to your reply.

\closing{Sincerely, \\ Anand Jain \\ University of Chicago 2021 \\ anandj@uchicago.edu}

\end{letter}

\end{document}