### 2020-03-16 random/career_preparation/cover_letters/wing.tex

\documentclass[11pt,a4paper]{letter} % 10pt font size default, 11pt and 12pt are also possible
\renewcommand*\familydefault{\sfdefault}

\address{610 Jackson St. \\ Santa Clara, CA 95050 \\ (408)597-4214 } % Your address and phone number
% \address{6103 S University Ave. \\ Chicago, IL 60637 \\ (408)597-4214 } % Your address and phone number

\begin{document}

\begin{letter}{Wing  \\ Palo Alto, CA 94304}

\opening{\textbf{Dear Wing,}}

I am interested in being an Engineering Intern for Summer 2020.

I believe I have the qualifications in programming, machine learning, and electronics to make a helpful contribution towards creating innovative applications of Wing planes.
Much of my time in college has been devoted towards thinking about neural network architectures and paradigms, differentiable programming and reinforcement learning being two primary interests. 
I am looking to further my understanding of the stack by working on embedded machine learning.

I have taken an electronics course, where we built radios, stopwatches, and other things in VDHL, as well as an interactive devices course that taught 3D printing and wearable device building. 
My ideal conversation with an employee would be about autonomy in a multi-agent domain, specifically how to solve the following problem: 
At UChicago, we have a problem in the winter, many sidewalks need to be cleared and salted on short notice. 
The current solution is tractors and manually dispersing the salt. 
I would like to see planes map unwalkable locations and report this to people cleaning up snow.

Thank you for your time and consideration.

I look forward to your reply.

\closing{Sincerely, \\ Anand Jain \\ University of Chicago 2021 \\ anandj@uchicago.edu}

\end{letter}

\end{document}