### 2020-04-02 random/career_preparation/cover_letters/intel_data.tex
bread.md and paper

\documentclass[11pt]{letter} % 10pt font size default, 11pt and 12pt are also possible
\renewcommand*\familydefault{\sfdefault} 
\usepackage[margin=1in]{geometry}

\address{610 Jackson St. \\ Santa Clara, CA 95050 \\ (408)597-4214 } % Your address and phone number
\address{6103 S University Ave. \\ Chicago, IL 60637 \\ (408)597-4214 } % Your address and phone number

\begin{document}

\begin{letter}{}

\opening{\textbf{Dear Intel,}}

\begin{flushright}
    Sincerely, \\ Anand Jain \\ University of Chicago 2021 \\ anandj@uchicago.edu
\end{flushright}

\end{letter}

\end{document}